---
title: "visualize-ecoregion-ch4-flux"
author: "Gavin McNicol"
date: "2/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Purpose

Code to visualize global wetland methane fluxes (TgCH4) based upon a 1) paint-by-numbers approach which assumes towers are representative of their constituency and 2) the UpCH4 predictions which does not make that assumption but which may be the case by chance.

## Packages

```{r load-packages, message = F}
rm(list=ls())
library(tidyverse) 
library(raster)
library(ncdf4)
library(sf)
library(rworldmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggthemes)
source("ggplot_theme.R")
```

## Get data

```{r get-data}
fluxes <- read_csv("/Volumes/Samsung_T5/Stanford/upch4_local/training-data/final.csv")
upch4_mean <- read_rds("/Volumes/Samsung_T5/Stanford/upch4_local/predictions/gridded/tg/global-regional/upch4_mean_stack.rds")[[2]]
upch4_sd <- read_rds("/Volumes/Samsung_T5/Stanford/upch4_local/predictions/gridded/tg/global-regional/upch4_sd_stack.rds")[[2]]
regions <- read_csv("/Volumes/Samsung_T5/Stanford/upch4_local/dissimilarity/output/stats/all_monthly_global_DI_subset.csv") %>% mutate(region = as.numeric(factor(closest_fold)))
wetlands <- stack("/Volumes/Samsung_T5/Stanford/upch4_local/grids/WAD2M/gcp-ch4_wetlands_2000-2018_025deg.nc")
```

## Create mask from x-y points

```{r}
xy <- list()
x <- list()
for(i in 1:26){
  xy[[i]] <- x
}

for(i in 1:26){
  for(j in 1:12){
  
x <- regions %>% 
  filter(region == i,
         Month == j) %>% 
  dplyr::select(x,y) %>% 
  distinct() %>% 
  dplyr::select(x) %>% 
  pull()

y <- regions %>% 
  filter(region == i,
         Month == j) %>% 
  dplyr::select(x,y) %>% 
  distinct() %>% 
  dplyr::select(y) %>% 
  pull()
  

xy[[i]][[j]] <- cbind(x, y)


# z <- regions %>% 
#   filter(Month == i) %>% 
#   dplyr::select(region) 
#   
# r <- raster(ncols = 360*4, nrows = 180*4)
# crs(r) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" 
#   
# region_mask[[i]] <- rasterize(xy, r, field = z)
#   
} 
}
# 
# plot(upch4_mean[[8]])
# points(xy[[1]][[8]])
```

## Get UpCH4 TD predictions for each constituency

```{r}
region_month_tg <- list()
x <- list()
for(i in 1:26){
  region_month_tg[[i]] <- x
}

year_vector <- rep(2001:2018, each = 12)
month_vector <- rep(1:12, 18)

for (i in 1:26){
  for (j in 1:216){
    
    if (length(xy[[i]][[ month_vector[j] ]]) == 0) {
      next
    }
  
    region_month_tg[[i]][[j]] <-
      raster::extract(upch4_mean[[j]], xy[[i]][[ month_vector[j] ]]) %>%
      as_tibble() %>%
      # pivot_longer(1:216,
      #              names_to = "timestep",
      #              values_to = "fch4_tg") %>%
      # mutate(timestep = as.numeric(str_extract(timestep, '\\b\\w+$'))) %>%
      # group_by(timestep) %>%
      summarize(fch4_tg = sum(value, na.rm = T)) %>% 
      mutate(region = i,
             month = month_vector[j],
             year = year_vector[j]) %>% 
      dplyr::select(region, year, month, fch4_tg)
    
  }
}

upch4_region_month_tg <- bind_rows(region_month_tg) 

# extract_tg(stack = upch4_mean, xy[[1]])
# region_tg <- lapply(xy, extract_tg, stack = upch4_mean) 
```

## Get 2000-2018 monthly wetland area for each constituency

```{r}
region_month_wa <- list()
x <- list()
for(i in 1:26){
  region_month_wa[[i]] <- x
}

year_vector <- rep(2000:2018, each = 12)
month_vector <- rep(1:12, 19)

for (i in 1:26){
  for (j in 1:228){
    
    if (length(xy[[i]][[ month_vector[j] ]]) == 0) {
      next
    }

    region_month_wa[[i]][[j]] <-
      raster::extract(wetlands[[j]] * area(wetlands[[j]]) * 10^6, xy[[i]][[ month_vector[j] ]] ) %>%
      as_tibble() %>%
      summarize(wetlands_m2 = sum(value, na.rm = T)) %>% 
      mutate(region = i,
             month = month_vector[j],
             year = year_vector[j]) %>% 
      dplyr::select(region, year, month, wetlands_m2)
    
  }
}

region_month_wa <- bind_rows(region_month_wa) 

# just get the monthly averages
region_month_wa <- region_month_wa %>% 
  group_by(region, month) %>% 
  summarize(wetlands_m2 = mean(wetlands_m2))
```

## Compute region-mean flux

```{r compute-mean-by-region}
tower_fch4_tg <- fluxes %>% 
  group_by(Cluster, Month) %>% 
  rename(region = Cluster, month = Month) %>% 
  summarize(region_mean_fch4 = mean(FCH4, na.rm = T)) %>% 
    left_join(region_month_wa) %>% 
  mutate(region_mean_fch4_tg = region_mean_fch4 * 16.04 * 10^-9 * 10^-12 * (60*60*24*30) * wetlands_m2) %>% 
  group_by(region) %>% 
  summarize(tower_fch4_tg = sum(region_mean_fch4_tg))
```

## Summarize results

```{r}
region_sums <- upch4_region_month_tg %>% 
  group_by(region, month) %>% 
  summarize(mean_fch4_tg = mean(fch4_tg)) %>% 
  group_by(region) %>% 
  summarize(upch4_fch4_tg = sum(mean_fch4_tg))

region_names <- regions %>% 
  dplyr::select(region_id = closest_fold, region) %>% 
  distinct()

region_names %>% 
  left_join(region_sums) %>% 
  left_join(tower_fch4_tg) %>% 
  mutate(upch4_tower_diff = upch4_fch4_tg - tower_fch4_tg) %>% 
  arrange(desc(upch4_tower_diff)) %>% 
write_csv("/Volumes/Samsung_T5/Stanford/upch4_local/predictions/gridded/tg/paint-by-numbers/constituency-methane-sums.csv")
```

## Map August constituencies with same colors (max wetland area)

```{r}
# get colors
col_vector_values <- read_csv("/Volumes/Samsung_T5/Stanford/upch4_local/dissimilarity/figures/200827_color_ramp_constituency.csv") %>% 
  dplyr::select(color_id) %>% 
  pull()

# get constituency areas for August 
xy_subset <- list()
for(i in 1:26){
  xy_subset[[i]] <- xy[[i]][[8]] %>% 
    as_tibble() %>% 
    mutate(region = i)
}
xy_august <- bind_rows(xy_subset) 

# add difference in emissions
region_diff <- read_csv("/Volumes/Samsung_T5/Stanford/upch4_local/predictions/gridded/tg/paint-by-numbers/constituency-methane-sums.csv") %>% dplyr::select(region_id, region, upch4_tower_diff) %>% 
  mutate(region_id = paste0(substr(region_id, 1, 2), "-", substr(region_id, 3, 5))) %>% 
  arrange(desc(upch4_tower_diff)) %>% 
  mutate(order = 1:n(),
         alpha_set = ifelse(abs(upch4_tower_diff) > quantile(abs(upch4_tower_diff), 0.5), 1, 0.5))

# merge
xy_august <- xy_august %>% 
  left_join(region_diff)

# get coordinates for tower locations
xy_august_subset <- xy_august %>% 
  dplyr::select(region_id, order) %>% 
  distinct() 

sites <- read_csv("/Volumes/Samsung_T5/Stanford/upch4_local/fluxnet-ch4-data/metadata/fluxnet-ch4-site-metadata.csv") %>% 
  dplyr::select(region_id = ID, x = Longitude, y = Latitude) %>% 
  right_join(xy_august_subset) 

## Make an outline of the countries for the plot
data(coastsCoarse)
map_outline_df <- fortify(coastsCoarse)
sPDF <- getMap()[getMap()$ADMIN!='Antarctica',]

abs(xy_august$upch4_tower_diff)/quantile(abs(xy_august$upch4_tower_diff), 0.1)

# constituency plots
ggplot() +
  geom_polygon(data= sPDF, aes(long, lat, group=group), fill = "grey90") +
  geom_tile(data = xy_august, aes(x = x, y = y, fill = fct_reorder(region_id, order), 
                                  alpha = alpha_set)) +
  geom_point(data = sites,
             aes(x = x, y = y,
                 color = fct_reorder(region_id, order)), 
             size = 1,
             shape = 21, 
             color = "black") +
  geom_path(data = map_outline_df, aes(long, lat, group = group), color = 'grey20', size = 0.3) +
  coord_equal() +
  scale_fill_manual(values = rev(col_vector_values[1:26])) +
  # scale_color_manual(values = rev(col_vector_values[1:26])) +
  scale_y_continuous(limits = c(-60, 85)) +
  labs(title = "Tower Constituency") +
  theme_map(8) +
  theme(legend.position = 'right')
ggsave("/Volumes/Samsung_T5/Stanford/upch4_local/predictions/gridded/tg/paint-by-numbers/constituency_tower_rep.png", height = 4.5, width = 8)

```

## Noticed that FR-LGt extrapolates a bit strangely, check differnces inputs

```{r}
data <- read_csv("/Volumes/Samsung_T5/Stanford/upch4_local/training-data/final.csv")

feat <- read_csv("/Volumes/Samsung_T5/Stanford/upch4_local/ffs/stepwise/feat_best_row_final.csv") %>%
  dplyr::select(feat) %>% pull() %>% as.character()
feat <- c("TA", "canopyht", feat[2:length(feat)])
feat <- feat[1:6] # select best
feat_l <- length(feat)

data %>% 
  dplyr::select(ID, feat) %>% 
  group_by(ID) %>% 
  summarize_all(.funs = mean) %>% 
  arrange(wc_mtdq)


```




